Digital Twins for Ocean Robots

### 2024-11-12 paper/paper.tex
update refs

The model hierarchy includes fast climate model emulators, global model output, ocean reanalyses, high-resolution models, and several ways to represent marine ecosystems. Through a streamlined workflow, \pkg{ClimateModels.jl} (JuliaCon 2021, 2023) makes it easy to operate these models that can provide a virtual environment for the virtual ocean robot fleet to observe and navigate. \pkg{IndividualDisplacements.jl} \cite{Forget2021} can be use to predict pathways of ocean robots that tend to follow ocean currents. \pkg{MeshArrays.jl} adds basic geospatial support for global climate model grids, providing capabilities such as interpolation and geolocation on a grid. The framework leverages and links to a variety of highly capable Julia packages from the community. It notably provides extension for visualization with \pkg{Makie.jl}, and tutorial examples in the form of \pkg{Pluto.jl} notebooks.

@@

\begin{table}
\tbl{Ocean observing platforms targeted by \pkg{OceanRobots.jl}, and associated data structures (or blank if not yet implemented).}{
\begin{tabular}{|l||r|}\hline
Platform Type & Data Structure\\\hline\hline
surface drifters & \struct{SurfaceDrifter}, \struct{CloudDrift} \\
drifting profilers & \struct{ArgoFloat} \\
moored buoys & \struct{OceanSite}, \struct{NOAAbuoy} \\
sea gliders & \struct{Gliders} \\
sail drones & \tbd \\
research vessels & \tbd \\
ships of opportunity & \tbd \\
orbiting satellites & \struct{SeaLevelAnomaly} \\
suborbital flights & \tbd \\
marine mammals & \tbd \\\hline
\end{tabular}}
\label{tab:robots}
\end{table}

@@

files_list=GDAC.files_list(); wmo=6900900
argo=read(ArgoFloat(),wmo=wmo,files_list=lst)

fig=plot(argo,option=:standard,pol=pol)
\end{lstlisting}

@@

Here we used the OCCA climatology \cite{Forget2010} to provide $T,S(x,y,z,t)$ and create Fig.~\ref{fig:Argo_Float_simu}, with \pkg{MeshArrays.jl} providing the geospatial tools. OCCA is essentially a previously trained model that consists of 12 monthly three-dimensional fields of $T,S$ (one per calendar month). Let's note that Figs.~\ref{fig:Argo_Float} and~\ref{fig:Argo_Float_simu} show broadly similar patterns -- this reflects that the predictive model for $T,S$ is skillful and that $T,S$ contrasts see in Fig.~\ref{fig:Argo_Float} largely reflect the $T,S$ sensor moving with the Argo float across the Ocean's geography. Differences between Figs.~\ref{fig:Argo_Float} and~\ref{fig:Argo_Float_simu} can in turn provide a basis for improving the predictive model, for example by including time variability beyond the seasonal cycle, or through various methods for data assimilation, artificial intelligence, or parameter inference \cite{Forget2010,Forget2015a,FFL15,FP15,Forget2024b}.

@@

Our second example focuses on the simulation of observing platforms moving with ocean currents. This basic simulation of surface drifter pathways uses the geostatistical average of velocities calculated in Fig.~\ref{fig:drifters_ve} as a predictive model for the $U,V(x,y,t)$ velocity fields. This gridded data set is simply provided as input to \pkg{IndividualDisplacements.jl}, which then releases virtual drifters and calculates their trajectories following the flow field (code~\ref{lst:IndDispCode}). The combination of these two model elements represents a previously trained model to predict surface drifter trajectories. While it neglects time-variability and small scales in ocean currents altogether, this model is sufficient to capture the mean pathway of sea water through this region -- the well-known loop current feeding the Gulf Stream via Florida Strait (Fig.~\ref{fig:drifters_speed_simu}).

@@

\begin{figure}[t]
\centerline{\includegraphics[width=\columnwidth]{figs/20240529_speed_subset_IndDisp.png}}
\caption{Virtual drifter trajectories predicted using (1) just the climatological mean flow field estimated from drifters (Fig.~\ref{fig:drifters_ve}, and the northward component) and (2) \pkg{IndividualDisplacements.jl} to calculate trajectories.}
\label{fig:drifters_speed_simu}
\end{figure}

\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:IndDispCode},
    caption={Simulate surface drifter pathways in the Gulf of Mexico (Fig.~\ref{fig:drifters_speed}) from flow fields $u,v$ and initial positions $x,y$.}
]
using IndividualDisplacements

@@

The core of DTOR is formed by \pkg{OceanRobots.jl}, \pkg{ArgoData.jl}, and \pkg{MeshArrays.jl} for the physical twins, along with \pkg{ClimateModels.jl}, \pkg{MITgcm.jl}, and \pkg{IndividualDisplacements.jl} for the virtual twins (i.e., predictive modeling). The \pkg{ClimateModels.jl} interface streamlines the use of models implemented in various languages. Our current model hierarchy supported via \pkg{ClimateModels.jl} includes, but is not limited to, the models used in this paper. DTOR can notably simulate observations in the future based on climate model predictions using \pkg{ClimateModels.jl}, which provides two options for this kind of applications -- either querying the CMIP6 archive of model output \cite{CMIP6} or using a fast emulator such as Hector \cite{Hector2015} (Fig.~\ref{fig:globalwarming}).

Within our model hierarchy, future climate scenarios like Fig.~\ref{fig:globalwarming} can be downscaled using gridded climatologies and reanalyses \cite{Forget2010,Forget2015a,Forget2024a}. The ECCO4 and OCCA2 reanalyses are simple to rerun with perturbed surface forcing fields via \pkg{MITgcm.jl}, which is partiucularly convenient for such applications. Multi-decadal solutions produced in this way can then be further downscaled using km-scale model output (Fig.~\ref{fig:km_scale}). At the end of this modeling workflow, \pkg{OceanRobots.jl}, \pkg{MeshArrays.jl} and \pkg{IndividualDisplacements.jl} enable calculations such as Figs.~\ref{fig:Argo_Float_simu} and \ref{fig:drifters_speed_simu}, for any of our models. DTOR can also take advantage of gridded satellite data, incl. sea surface temperature and sea level anomalies, and it is sometimes possible to use these instead of ocean reanalyses.

@@

Our current focus is on geospatial analyses that track global warming and marine heatwaves using Argo data \cite{Forget2024a}. Another example from our research is with the CStream research program (UK-US) which is releasing drifters near Florida Strait to track the nutrient stream that is associated with the Gulf Stream (Fig.~\ref{fig:Cstream}). Other applications at the scale of oceanic basins include tracking ocean plastic pollution, monitoring biological impacts of extreme events and global warming, or the optimization of the global climate monitoring fleet of ocean robots for e.g. cost. More locally, DTOR could be used in the context of field experiments such as S-MODE, to choose an example from the recent past. Related projects on digital twins, not focused on exploiting Julia, include the EU funded {\it Destination Earth} program, Mercator Ocean International's {\it Digital Twin Ocean}, and the UN Decade Action's DITTO initiative ({\it Digital Twins of The Ocean}).

>>>>

The model hierarchy includes fast climate model emulators, global model output, ocean reanalyses, high-resolution models, and several ways to represent marine ecosystems. Through a streamlined workflow, \pkg{ClimateModels.jl} (JuliaCon 2021, 2023) makes it easy to operate these models that can provide a virtual environment for the virtual ocean robot fleet to observe and navigate. \pkg{Drifters.jl} \cite{Forget2021} can be use to predict pathways of ocean robots that tend to follow ocean currents. \pkg{MeshArrays.jl} adds basic geospatial support for global climate model grids, providing capabilities such as interpolation and geolocation on a grid. The framework leverages and links to a variety of highly capable Julia packages from the community. It notably provides extension for visualization with \pkg{Makie.jl}, and tutorial examples in the form of \pkg{Pluto.jl} notebooks.

@@

\begin{table}
\tbl{Ocean observing platforms targeted by \pkg{OceanRobots.jl}, and associated data structures (or blank if not yet implemented).}{
\begin{tabular}{|l||r|}\hline
Platform Type & Data Structure\\\hline\hline
surface drifters & \struct{SurfaceDrifter}, \struct{CloudDrift} \\
drifting profilers & \struct{ArgoFloat} \\
moored buoys & \struct{OceanSite}, \struct{NOAAbuoy} \\
sea gliders & \struct{Gliders} \\
expendable bathythermographs & \tbd \\
sea gliders & \struct{Gliders} \\
sail drones & \tbd \\
research vessels & CCHDO \\
ships of opportunity & \tbd \\
orbiting satellites & \struct{SeaLevelAnomaly} \\
suborbital flights & \tbd \\
marine mammals & \tbd \\\hline
\end{tabular}}
\label{tab:robots}
\end{table}

@@

lst=GDAC.files_list(); wmo=6900900
argo=read(ArgoFloat(),wmo=wmo,files_list=lst)

fig=plot(argo,option=:standard)
\end{lstlisting}

@@

To create Fig.~\ref{fig:Argo_Float_simu}, we use \pkg{Climatology.jl} to access the OCCA climatology \cite{Forget2010}, and then interpolate $T,S(x,y,z,t)$ via the geospatial tools in \pkg{MeshArrays.jl} providing. OCCA is essentially a previously trained model that consists of 12 monthly three-dimensional fields of $T,S$ (one per calendar month). Let's note that Figs.~\ref{fig:Argo_Float} and~\ref{fig:Argo_Float_simu} show broadly similar patterns -- this reflects that the predictive model for $T,S$ is skillful and that $T,S$ contrasts see in Fig.~\ref{fig:Argo_Float} largely reflect the $T,S$ sensor moving with the Argo float across the Ocean's geography. Differences between Figs.~\ref{fig:Argo_Float} and~\ref{fig:Argo_Float_simu} can in turn provide a basis for improving the predictive model, for example by including time variability beyond the seasonal cycle, or through various methods for data assimilation, artificial intelligence, or parameter inference \cite{Forget2010,Forget2015a,FFL15,FP15,Forget2024b}.

@@

Our second example focuses on the simulation of observing platforms moving with ocean currents. This basic simulation of surface drifter pathways uses the geostatistical average of velocities calculated in Fig.~\ref{fig:drifters_ve} as a predictive model for the $U,V(x,y,t)$ velocity fields. This gridded data set is simply provided as input to \pkg{Drifters.jl}, which then releases virtual drifters and calculates their trajectories following the flow field (code~\ref{lst:IndDispCode}). The combination of these two model elements represents a previously trained model to predict surface drifter trajectories. While it neglects time-variability and small scales in ocean currents altogether, this model is sufficient to capture the mean pathway of sea water through this region -- the well-known loop current feeding the Gulf Stream via Florida Strait (Fig.~\ref{fig:drifters_speed_simu}).

@@

\begin{figure}[t]
\centerline{\includegraphics[width=\columnwidth]{figs/20240529_speed_subset_IndDisp.png}}
\caption{Virtual drifter trajectories predicted using (1) just the climatological mean flow field estimated from drifters (Fig.~\ref{fig:drifters_ve}, and the northward component) and (2) \pkg{Drifters.jl} to calculate trajectories.}
\label{fig:drifters_speed_simu}
\end{figure}

\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:IndDispCode},
    caption={Simulate surface drifter pathways in the Gulf of Mexico (Fig.~\ref{fig:drifters_speed}) from flow fields $u,v$ and initial positions $x,y$.}
]
using Drifters

@@

The core of DTOR is formed by \pkg{OceanRobots.jl}, \pkg{ArgoData.jl}, and \pkg{MeshArrays.jl} for the physical twins, along with \pkg{ClimateModels.jl}, \pkg{MITgcm.jl}, and \pkg{Drifters.jl} for the virtual twins (i.e., predictive modeling). The \pkg{ClimateModels.jl} interface streamlines the use of models implemented in various languages. Our current model hierarchy supported via \pkg{ClimateModels.jl} includes, but is not limited to, the models used in this paper. DTOR can notably simulate observations in the future based on climate model predictions using \pkg{ClimateModels.jl}, which provides two options for this kind of applications -- either querying the CMIP6 archive of model output \cite{CMIP6} or using a fast emulator such as Hector \cite{Hector2015} (Fig.~\ref{fig:globalwarming}).

Within our model hierarchy, future climate scenarios like Fig.~\ref{fig:globalwarming} can be downscaled using gridded climatologies and reanalyses \cite{Forget2010,Forget2015a,Forget2024a}. The ECCO4 and OCCA2 reanalyses are simple to rerun with perturbed surface forcing fields via \pkg{MITgcm.jl}, which is partiucularly convenient for such applications. Multi-decadal solutions produced in this way can then be further downscaled using km-scale model output (Fig.~\ref{fig:km_scale}). At the end of this modeling workflow, \pkg{OceanRobots.jl}, \pkg{MeshArrays.jl} and \pkg{Drifters.jl} enable calculations such as Figs.~\ref{fig:Argo_Float_simu} and \ref{fig:drifters_speed_simu}, for any of our models. DTOR can also take advantage of gridded satellite data, incl. sea surface temperature and sea level anomalies, and it is sometimes possible to use these instead of ocean reanalyses.

@@

Our current focus is on geospatial analyses that track global warming and marine heatwaves using Argo data \cite{Forget2024a}. Another example from our research is with the C-Streams research program (UK-US) which is releasing drifters near Florida Strait to track the nutrient stream that is associated with the Gulf Stream. The example in Fig.~\ref{fig:Cstream} is based on the monthly mean ECCO climatology for ocean transports \cite{Forget2015a,Forget2019,Rousselet2021}, and using \pkg{Climatology.jl}, \pkg{MeshArrays.jl}, and \pkg{Drifters.jl} to calculate pathways. Other applications at the scale of oceanic basins include tracking ocean plastic pollution, monitoring biological impacts of extreme events and global warming, or the optimization of the global climate monitoring fleet of ocean robots for e.g. cost. More locally, DTOR could be used in the context of field experiments such as S-MODE, to choose an example from the recent past. Related projects on digital twins, not focused on exploiting Julia, include the EU funded {\it Destination Earth} program, Mercator Ocean International's {\it Digital Twin Ocean}, and the UN Decade Action's DITTO initiative ({\it Digital Twins of The Ocean}).
